### 2016-08-09 ms/ms.tex
finsihed first draft of introduction

Keywords: changing points, temporal networks, hedgerows, species
interactions, preferential attachment

@@

Global change has created a severe biodiversity crisis
\citep{barnosky2011has}. As species are lost, so are their
interactions, causing further coextinction cascades
\citep{dunn2009sixth}.  Mutualistic systems are particualrly at risk.
Ecological restoration plays an increasingly important role in
recovering biodiversity (CITATIONS). Because the structure of
interaction networks is related to ability of communities to maintain
function in the face of species extinction \citep{Memmott2004}, a key
restoration aim is to facilitate assembly of robust networks. Little
is known, however, about how restoration influences the assembly of
interaction networks, or ecological network assembly more generally.

To date, only two field studies have examined how networks assemble
over time, often using space for time gradients. \cite{Olesen2008} was
investigated day-to-day, temporal assembly of a plant-pollinator
network within a season, taking advantage of the extreme seasonality
of pollinator communities in Greenland.  \cite{Olesen2008} found that
within a season, the network assembly was similar to preferential
attachment. New species tended to interact with already well-connected
species, likely because these species are either more abundant or more
temporally persistent.

The mostly widely explored mechanism of network assembly, preferential
attachment \citep{barabasi1999emergence}, predicts that a new species
is more likely to interact with species that are already
well-connected \citep[''the rich-get-richer''
principle,][]{barabasi1999emergence}.

\item Studying primary succession along a glacier foreland,
  \cite{albrecht2010plant} found a similar pattern where nestedness, a
  pattern of interactions where a generalist core interacts with both
  specialist and generalist species, increased as the community
  aged. % Studying ``managed succession'' of a clear-cut pine forest,
  % \cite{devoto2012understanding} found changes in network structure
  % were explained by a combination of age, tree density and variation
  % in tree diameter. 
\item Even non-successional temporal dynamics suggest a stable core of
  generalists persist despite high turnover of peripheral species
  \citep{fang2012relative, diaz2010changes, alarcon2008year}.
\item In contrast to the ordered network build-up described by
  preferential attachment, assembly may be punctuated by significant
  reorganizations of interactions. For example, as new species are
  added, resident species change their interaction partners to
  minimize competition, or become extinct. Such significant
  reorganizations of interactions, or changing points, have been
  observed in networks \citep{peel2014detecting}.
\end{itemize}

@@

Using Bayesian posterior inference and techniques from phylogenetic
tree reconstruction, we fit the GHRG model to the networks
\citep{peel2014detecting}. This is accomplished by using a Markov
Chain Monte Carlo (MCMC) procedure to first sample the posterior
distribution of bipartitions, from which a consensus tree is derived
\citep{peel2014detecting}. ADD NEWMAN PAPER. We used $\beta$
distributions with the hyperparameters $\alpha=\beta=1$ to define
priors for $p_r$.

@@

% \subsubsection*{Characteristics of species that contribute to change
%   points}

% To further elucidate the nature of the change points, we examine the
% characteristics of the species that contributed the reorganization of
% interactions. Some species remain in relativly similar network
% positions through time, whereas others are more variable in their role
% and thus contribute more strongly to network reorganization. If
% assembly is following according to a mechanism of preferential
% attachment, the more persistnet species with the higest degree would
% be most stable in their network position ???

% We test the hypothesis that the species that occupy the most stable
% positions in the network are the most temporally persistent and or
% generalized \citep{fang2012relative}.

% We calculate species persistence as the proportion of the surveys a
% pollinator was observed. Species observed consistently within and
% between years are thus maximally persistent. Species degree is based
% on plant-pollinator interaction observations from a more extensive
% data-set from Yolo County (approx.~ 18000 interaction records) that
% included both the data included in this study and additional data from
% sites where we collected flower visitors using the same methods
% \citep{mgonigle-2015-x, ponisio2015farm}.

\subsection*{Temporal changes in interaction patterns}

The changing points in network structure may contribute to the
reorganization of the assembling networks into predictable interaction
patterns. Pollination networks exhibit two main topologies ---
modularity \citep[e.g.,][]{Olesen2007} and nestedness
\citep[e.g.,][]{Bascompte2006, Bascompte2003}. Modular community
interactions are insular, occurring within separate groups or
``modules'' more often than between modules. Communities in the
network may fragment as the network assembles, enhancing
modularity. Conversely, nested networks are like pyramid of
interactions, where there are some species that interact with many
species, other species that interact with a subset of those species,
and so on. If species entering the network tend to interact with the
generalist base of the network pyramid (i.e., preferential
attachment), nestedness would increase through time. Lastly, if the
network is accumulating specialist species or species are beginning to
limit their interaction niche breath as the network assembles, this
would lead to an increase in the network-level specialization
\citep{bluthgen-2006-9}. To test whether network modularity,
nestedness or specialization changed linearly with assembly, we used
linear mixed models with the descriptive network metrics as the
response variable, year of assembly as the explanatory variable, and
random effects of site and year. We use NODF (weighted or unweighted)
to evaluate network nestedness \citep{nodf}. NODF evaluates whether
species with fewer partners interact with subsets of partners with
which more connected species interact \citep{nodf}. We used a
hierarchical clustering algorithm to estimate modularity
\citep{Newman2004, igraph}.

We calculated standardized $z$-scores so that network metrics could be
compared between communities. The $z$-scores were calculated by
generating an ensemble of $999$ randomly assembled communities,
subtracting the mean of the statistic calculated across these
communities from the observed value, and then dividing by the standard
deviation. To assemble random communities, we reshuffled the
interactions between species but fixed the total number of
interactions, species and the distribution of the interaction
frequencies \citep{Galeano2009}.

\subsection*{Core and peripherial communities}

@@

Lastly, we test the changes in interaction patterns associated with
network assembly affect the robustness of the network to species
loss. Following \cite{Memmott2004}, we simulate the extinction of
plant species the subsequent extinction cascades of pollinator
species. Because the reproduction of plant species if facilitated by
active restoration efforts, it is unlikely the extinction of
pollinators would effect plant populations in the hedgerows. However,
plants ceasing to bloom, for example in response to drought, will
likely effect the pollinators that depend on them. Plants species were
eliminated based on their degree or abundance, and the number of
pollinators that secondarily went extinct is calculated. The area
below the extinction curve is used as a measure of network robustness.

% \subsubsection*{Characteristics of ``core'' and ``peripheral''
%   communities}

% \begin{figure}
%   \centering
%   \includegraphics[width=.8\textwidth]{../analysis/changePoint/plotting/communities.pdf}
%   \caption{XXX}
%   \label{fig:communities}
% \end{figure}
% \clearpage

\begin{figure}
  \centering
  \includegraphics[width=.7\textwidth]{../analysis/changePoint/plotting/figures/pcoa/pcoa.pdf}
  \caption{XXX}
  \label{fig:pcoa}
\end{figure}
\clearpage

%% no need for robustness figs bec not sig

We would like to thank Leto Peel and Aaron Clauset for their
invaluable discussions and for help with the change point analysis.
We thank the growers and land owners that allowed us to work on their
property.  We also appreciate the identification assistance of expert
taxonomists Martin Hauser, Robbin Thorp and Jason Gibbs.  This work
was supported by funding from the Army Research Office
(W911NF-11-1-0361 to CK), the Natural Resources Conservation Service
(CIG-69-3A75-12-253, CIG-69-3A75-9-142, CIG-68-9104-6-101 and
WLF-69-7482-6-277 to The Xerces Society), the National Science
Foundation (DEB-0919128 to CK), The U.S.  Department of Agriculture
(USDA-NIFA 2012-51181-20105 to Michigan State University).  Funding
for LCP was provided by an NSF Graduate Research Fellowship and the
USDA NIFA Graduate Fellowship. FUNDING FOR MARILLIA.

>>>>

Keywords: changing points, temporal networks, hedgerows, species
interactions, preferential attachment, mutualisms

@@

%% introduce mutualisms, restoration, network assembly
Global change has created a severe biodiversity crisis, and as species
are lost, so are their interactions \citep{dunn2009sixth,
  barnosky2011has}. Because mutualistic interactions are essential for
maintaining the diversity their component guilds of species, these
systems are particularly at risk from coextinction cascades. The
nature of these coextinction cascades depends on the interaction
patterns within a community \citep{Memmott2004, Rezende2007,
  Bascompte2009}. Recovering the lost biodiversity and interactions
through ecological restoration has become increasingly imperative, and
a key restoration aim is to facilitate assembly of robust interaction
networks \citep{menz-2010-4}. We know little, however, about how to
re-assemble interacting communities through restoration, or the
process of ecological network assembly more generally.

%% most people think network assemble via preferential attachment, and
%% there is some evidence of that. (other studies on mutualisms?)
The mostly widely explored mechanism of network assembly, preferential
attachment \citep{barabasi1999emergence}, predicts that a new species
is more likely to interact with species that are already
well-connected \citep[''the rich-get-richer''
principle,][]{barabasi1999emergence}. In pollination systems --- a
particularly ubiquitous mutualism \citep{ollerton-2011-321,
  klein-2007-303} --- some studies have found support for this
mechanism of assembly. Investigating the day-to-day, temporal assembly
of a plant-pollinator network within a season, \cite{Olesen2008} found
that new species tended to interact with already well-connected
species, likely because these species are either more abundant or more
temporally persistent \citep{Olesen2008}. In addition, using a
space-for-time substitution to study primary succession along a
glacier foreland, \cite{albrecht2010plant} found some indication
assembly was occurring through preferential attachment. Network
nestedness, a pattern of interactions where a generalist core
interacts with both specialist and generalist species, increased as
the community aged \citep{albrecht2010plant}. Increasing nestedness
could result from a process like preferential attachment where
specialist species attach to the well-connected, generalist core.

%% there are other ways to assemble that are not preferentia;
%% attchment, particularly vis big reorganizations of interactions
In contrast to the ordered network build-up described by preferential
attachment, assembly can be punctuated by significant reorganizations
of interactions \citep{peel2014detecting}. For example, as new species
are added, resident species change their interaction partners to
minimize competition, or become extinct. Such significant
reorganizations of interactions, or changing points, have been
observed in networks responding to abrupt shifts in the behavior of
interacters \citep{peel2014detecting}. No studies, however, have
examined whether changing points occur during ecological network
assembly.

%% restoration
Understanding network assembly is particularly relevant to ecological
restoration, which is essentially 'applied succession'
\citep[e.g.,]{parker1997scale}. In pollination systems, the time since
an area was restored has been shown to effect the structure of
networks \citep{forup-2008-742, forup2008restoration,
  devoto2012understanding}, suggesting interactions are evolving as the
community develops. Understanding the mechanisms of network assembly
will help to guide the restoration of particular communities.

%% restoration of ecosystem services is really important, particularly
%% in ag. but ag is a bad place for pollinators
Facilitating effective restoration of networks is particularly
imperative in areas where species interactions provide essential
ecosystem services, such as crop pollination. In intensively managed
agricultural landscapes, the demand for pollination services is the
greatest \citep{kremen-2008-10}. However, honey bees, managed
extensively around the world to provide crop pollination, are in
global decline \citep{neumann-2010-1, van-engelsdorp-2009-e6481}. In
addition, native pollinators, which have the capacity to provide
sufficient crop pollination \citep{kremen-2002-16816,
  winfree-2007-1105, kremen-2004-1109}, are in short supply because
these landscapes make poor habitats for pollinator populations
\citep{kremen-2002-16816}. To ensure provision the continued provision
of ecosystem services and curb biodiversity loss, effective
restoration of pollinators and their interactions in agricultural
landscapes is critical.

%% hedgerows!!  
To promote pollinator services in agriculture, farmers are
increasingly turning to the habitat restoration technique of planting
strips of native plants along farm edges (hedgerows) to help provide
habitat for pollinators without removing arable land from
production. Hedgerows have been shown to augment the richness,
abundance and trait diversity of pollinators in agricultural
landscapes\citep{morandin-2013-829, mgonigle-2015-x, kremen-2015-602,
  ponisio2015farm}. In addition, hedgerows promote the persistence and
colonization of floral resource specialists
\citep{mgonigle-2015-x}. Little is known however, about the assembly
of the network following hedgerow restoration.

Using a long-term data-set of plant-pollinator communities assembling
following hedgerow restoration in the highly simplified and
intensively managed agricultural landscape of California's Central
Valley, we explore the process of network development. We first
determine whether the mechanism underlying network assembly is a
smooth build up of interactions as would be predicted by preferential
attachment, or punctuated by significant reorganizations of
interactions (i.e., changing points). Even with changing points in
interaction organization, networks could still be assembling via
preferential attachment if the network reorganizations were primarily
driven the by peripheral, temporally variable species while a stable,
well-connected core of species still persists. We thus examine whether
the species are most variable in their network position --- and thus
important contributors network reorganizations --- are less persistent
and connected species. %% interaction beta div, time beta div??
Lastly, we examine whether networks are assembling toward predictable
interaction patterns, and the ramifications for the robustness of the
networks to species extinction and perturbation.

%% objectives
% 1) Are hedgerows assembling by preferential attachment, or are there
% more abrupt reorganizations of network structure?
% 2) Who are the species responsible for interaction reorganizations?
% how to incorporate this? 
% 3) Are the changing points a product of turnover in species
% composition, or are more substantial changes in the verticles of the
% networks?
% 4) Is the structure of interactions changing? Does this change the
% robustness of the network to species extinction and/or perturbation?

% Pollinators are one such key group of interactors: $75\%$ of all
% crop species depend to some extent on pollinator visitation
% \citep{klein-2007-303}, and animal-pollinated crops supply a large
% proportion of essential nutrients to the human diet
% \citep{eilers-2011-e21363}.

% Even non-successional temporal dynamics suggest a stable core of
% generalists persist despite high turnover of peripheral species
% \citep{fang2012relative, diaz2010changes, alarcon2008year}.

@@

Using Bayesian posterior inference and techniques from phylogenetic
tree reconstruction, we fit the GHRG model to the networks
\citep{peel2014detecting}. This is accomplished by using a Markov
Chain Monte Carlo (MCMC) procedure to first sample the posterior
distribution of bipartitions, from which a consensus tree is derived
\citep{peel2014detecting}. We used $\beta$ distributions with the
hyperparameters $\alpha=\beta=1$ to define priors for $p_r$.

@@

To further elucidate the nature of the change points, we examine the
characteristics of the species that contributed the reorganization of
interactions. Some species remain in relatively similar network
positions through time, whereas others are more variable in their role
and thus contribute more strongly to network reorganization. We test
the that more persistent species with the highest degree are the most
stable in their network position, as would be expected is the networks
were assembling via preferential attachment.

We calculate species persistence as the proportion of the surveys a
pollinator was observed. Species observed consistently within and
between years are thus maximally persistent. Weighted species degree
is calculated from interaction observations observed in more extensive
data-set from Yolo County (approx.~ 18000 interaction records) that
included both the data included in this study and additional data from
sites where we collected flower visitors using the same methods
\citep{mgonigle-2015-x, ponisio2015farm}.%% does this make sense to
                                %% calculate from an independent data set?
To represent the the variability of species within networks, we
computed the coefficient of variation of weighted closeness at each
site through time. Closeness describes the centrality of a species in
the network by calculating path lengths to other vertices (species) in
the graph. We used linear mixed models to test whether the variability
of species closeness values was related to the persistence or degree
of that species. We included random effects for species, as well as
site. We focused on the pollinator species because the hedgerow
flowers are planted and thus are not directly assembling. Because
degree and persistence were strongly correlated, ($\rho = 0.84$,
$p$-value $<$ $2*10^{-16}$), each explanatory variable was included in
the model separately.

\subsection*{Temporal changes in interaction patterns}
The changing points in network structure may contribute to the
reorganization of the assembling networks into predictable interaction
patterns. Pollination networks exhibit two main topologies ---
modularity \citep[e.g.,][]{Olesen2007} and nestedness
\citep[e.g.,][]{Bascompte2006, Bascompte2003}. Modular community
interactions are insular, occurring within separate groups or
``modules'' more often than between modules. Communities in the
network may fragment as the network assembles, enhancing
modularity. Conversely, nested networks are like pyramid of
interactions, where there are some species that interact with many
species, other species that interact with a subset of those species,
and so on. If species entering the network tend to interact with the
generalist base of the network pyramid (i.e., preferential
attachment), nestedness would increase through time. Lastly, if the
network is accumulating specialist species or species are beginning to
limit their interaction niche breath as the network assembles, this
would lead to an increase in the network-level specialization
\citep{bluthgen-2006-9}. To test whether network modularity,
nestedness or specialization changed linearly with assembly, we used
linear mixed models with the descriptive network metrics as the
response variable, year of assembly as the explanatory variable, and
random effects of site and year.

We use NODF (weighted or unweighted) to evaluate network nestedness
\citep{nodf}. NODF evaluates whether species with fewer partners
interact with subsets of partners with which more connected species
interact \citep{nodf}. To estimate modularity, we use a hierarchical
clustering algorithm \citep{Newman2004, igraph}.  We calculated
standardized $z$-scores so that nestedness and modularity metrics
could be compared between communities. The $z$-scores were calculated
by generating an ensemble of $999$ randomly assembled communities,
subtracting the mean of the statistic calculated across these
communities from the observed value, and then dividing by the standard
deviation. To assemble random communities, we reshuffled the
interactions between species but fixed the total number of
interactions, species and the distribution of the interaction
frequencies \citep{Galeano2009}. Lastly, Network specialization was
measured using H2, which estimate the deviation of the observed
interaction frequency between plants and pollinators from a null
expectation where all partners interact in proportion to their
abundances \citep{bluthgen-2006-9}. It ranges from 0 for generalized
networks to 1 for specialized networks.

Lastly, we test the changes in interaction patterns associated with
network assembly affect the robustness. First, following
\cite{Memmott2004}, we simulate the extinction of plant species the
subsequent extinction cascades of pollinator species. Because the
reproduction of plant species if facilitated by active restoration
efforts, it is unlikely the extinction of pollinators would effect
plant populations in the hedgerows. However, plants ceasing to bloom,
for example in response to drought, will likely effect the pollinators
that depend on them. Plants species were eliminated based on their
degree or abundance, and the number of pollinators that secondarily
went extinct is calculated. The area below the extinction curve is
used as a measure of network robustness \citep{Memmott2004}.

%% Marilia add perturbation methods here

We would like to thank Leto Peel and Aaron Clauset for their
invaluable discussions and for help with the change point analysis.
We thank the growers and land owners that allowed us to work on their
property.  We also greatly appreciate the identification assistance of
expert taxonomists Martin Hauser, Robbin Thorp and Jason Gibbs.  This
work was supported by funding from the Army Research Office
(W911NF-11-1-0361 to CK), the Natural Resources Conservation Service
(CIG-69-3A75-12-253, CIG-69-3A75-9-142, CIG-68-9104-6-101 and
WLF-69-7482-6-277 to The Xerces Society), the National Science
Foundation (DEB-0919128 to CK), The U.S.  Department of Agriculture
(USDA-NIFA 2012-51181-20105 to Michigan State University).  Funding
for LCP was provided by an NSF Graduate Research Fellowship and the
USDA NIFA Graduate Fellowship. FUNDING FOR MARILLIA.

% \subsection*{Core and peripherial communities}

@@
